Data Prep

### 2024-02-14 scripts/00.0_dataPrep.qmd
fixed file names to use here::here()

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
```

@@

```{r}
base::load("movebankCredentials", "pw.Rda")
MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
rm(pw)
```

@@

```{r download-inpa}
inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
inpa <- methods::as(inpa, "data.frame")
inpa <- inpa %>%
  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
         year = as.numeric(lubridate::year(timestamp))) %>%
  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
write_feather(inpa, "data/dataPrep/inpa.feather")

@@

```{r download-ornitela}
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-09-15 11:59"
ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
write_feather(ornitela, "data/dataPrep/ornitela.feather")

@@

```{r load-inpa-ornitela}
inpa <- read_feather("data/dataPrep/inpa.feather")
ornitela <- read_feather("data/dataPrep/ornitela.feather")
```

@@

# Add the Nili_id
ww <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

@@

joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                  by = c("trackId" = "mb_id"))
nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
length(unique(joined$Nili_id))
# For any that didn't get Nili_ids assigned, make it the trackId
joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
joined <- joined %>%
  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                             TRUE ~ Nili_id))
length(unique(joined$Nili_id))
write_feather(joined, "data/dataPrep/joined.feather")

@@

```{r load-hospital}
joined <- read_feather("data/dataPrep/joined.feather")
periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
```

@@

write_feather(removed_periods, "data/dataPrep/removed_periods.feather")

@@

```{r}
removed_periods <- read_feather("data/dataPrep/removed_periods.feather")
```

```{r remove-capture-cage-getroosts}
#| eval: false
# Identify and remove capture dates using Marta's code --------------------
roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
save(roosts0, file = "data/dataPrep/roosts0.Rda")

@@

```{r}
# Load information about the capture sites
capture_sites <- read.csv("data/raw/capture_sites.csv")
load("data/dataPrep/roosts0.Rda")
AllCarmelDates <- read.csv("data/raw/all_captures_carmel_2010-2021.csv")
```

@@

write_feather(removed_captures, "data/dataPrep/removed_captures.feather")

@@

```{r attach-age-sex}
#| message: false
removed_captures <- read_feather("data/dataPrep/removed_captures.feather")
age_sex <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()
```

@@

write_feather(with_age_sex, "data/dataPrep/with_age_sex.feather")

@@

```{r}
mask <- sf::st_read("data/raw/CutOffRegion.kml")
with_age_sex <- read_feather("data/dataPrep/with_age_sex.feather")
```

@@

save(cleaned, file = "data/dataPrep/cleaned.Rda")

@@

# Separate the seasons -----------------------------
seasons_list <- with_seasons %>%
  group_by(seasonUnique) %>%
  group_split(.keep = T)
length(seasons_list) # should be 9
# extract the season names in case we need a separate vector at some point.
season_names <- map_chr(seasons_list, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
length(season_names) # should also be 9
save(seasons_list, file = "data/dataPrep/seasons_list.Rda")
save(season_names, file = "data/dataPrep/season_names.Rda")

@@

```{r}
load("data/dataPrep/seasons_list.Rda")
load("data/dataPrep/season_names.Rda")
```

@@

save(low_fix_rate_indivs, 
     file = "data/dataPrep/low_fix_rate_indivs.Rda")
# Don't need to save with_modes because we've already used this step to identify low fix rate individuals.

@@

```{r}
load("data/dataPrep/seasons_list.Rda") 
load("data/dataPrep/low_fix_rate_indivs.Rda")
```

```{r}
removed_lfr <- map2(seasons_list, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
save(removed_lfr, file = "data/dataPrep/removed_lfr.Rda")  
# tidy up 
rm(seasons_list) 
rm(low_fix_rate_indivs) 
rm(removed_lfr) 
gc()
```

@@

```{r}
load("data/dataPrep/removed_lfr.Rda")
```

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = "data/dataPrep/roosts.Rda")

@@

```{r}
load("data/dataPrep/removed_lfr.Rda")
load("data/dataPrep/season_names.Rda")
```

@@

save(removed_nighttime, file = "data/dataPrep/removed_nighttime.Rda")

@@

```{r}
load("data/dataPrep/removed_nighttime.Rda")
load("data/dataPrep/season_names.Rda")
```

@@

save(removed_northern, file = "data/dataPrep/removed_northern.Rda")

@@

```{r}
load("data/dataPrep/removed_northern.Rda")
```

@@

save(removed_lowppd, file = "data/dataPrep/removed_lowppd.Rda")

@@

```{r}
load("data/dataPrep/removed_lowppd.Rda")
load("data/dataPrep/season_names.Rda")
```

@@

save(removed_too_few_days, file = "data/dataPrep/removed_too_few_days.Rda")
save(season_durations, file = "data/dataPrep/season_durations.Rda")

@@

```{r}
load("data/dataPrep/removed_too_few_days.Rda")
```

```{r}
#| eval: false
# Get elevation rasters ---------------------------------------------------
sf <- map(removed_too_few_days, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
save(elevs_z10, file = "data/dataPrep/elevs_z10.Rda")
save(sf, file = "data/dataPrep/sf.Rda")
```

```{r}
load("data/dataPrep/elevs_z10.Rda")
load("data/dataPrep/sf.Rda")
```

@@

save(with_altitudes, file = "data/dataPrep/with_altitudes.Rda")

>>>>

```{r load-packages}
library(vultureUtils)
library(sf)
library(tidyverse)
library(move)
library(feather)
library(readxl)
library(elevatr)
library(here)
```

@@

```{r}
base::load(here::here("movebankCredentials", "pw.Rda"))
MB.LoginObject <- move::movebankLogin(username = "kaijagahm", password = pw)
rm(pw)
```

@@

```{r download-inpa}
inpa <- move::getMovebankData(study = 6071688, login = MB.LoginObject, removeDuplicatedTimestamps = TRUE)
inpa <- methods::as(inpa, "data.frame")
inpa <- inpa %>%
  mutate(dateOnly = lubridate::ymd(substr(timestamp, 1, 10)),
         year = as.numeric(lubridate::year(timestamp))) %>%
  filter(lubridate::ymd(dateOnly) >= lubridate::ymd("2020-09-01"), lubridate::ymd(dateOnly) <= lubridate::ymd("2023-09-15")) # cut this off at the same point as the ornitela data
write_feather(inpa, here::here("data", "dataPrep", "inpa.feather"))

@@

```{r download-ornitela}
# Ornitela: download data from movebank (just a subset of the times for now)
minDate <- "2020-09-01 00:00"
maxDate <- "2023-09-15 11:59"
ornitela <- vultureUtils::downloadVultures(loginObject = MB.LoginObject, removeDup = T, dfConvert = T, quiet = T, dateTimeStartUTC = minDate, dateTimeEndUTC = maxDate)
write_feather(ornitela, here::here("data", "dataPrep", "ornitela.feather"))

@@

```{r load-inpa-ornitela}
inpa <- read_feather(here::here("data", "dataPrep", "inpa.feather"))
ornitela <- read_feather(here::here("data", "dataPrep", "ornitela.feather"))
```

@@

# Add the Nili_id
ww <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, Movebank_id) %>%
  distinct()

@@

joined <- left_join(joined0, ww %>% dplyr::select(mb_id, Nili_id) %>% distinct(),
                  by = c("trackId" = "mb_id"))
nrow(joined) == nrow(joined0) # good, same number of rows. (check that the join worked)
length(unique(joined$Nili_id))
# For any that didn't get Nili_ids assigned, make it the trackId
joined %>% filter(is.na(Nili_id)) %>% pull(trackId) %>% unique() # as expected.
joined <- joined %>%
  mutate(Nili_id = case_when(is.na(Nili_id) ~ trackId,
                             TRUE ~ Nili_id))
length(unique(joined$Nili_id))
write_feather(joined, here::here("data", "dataPrep", "joined.feather"))

@@

```{r load-hospital}
joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
periods_to_remove <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
```

@@

write_feather(removed_periods, here::here("data", "dataPrep", "removed_periods.feather"))

@@

```{r}
removed_periods <- read_feather(here::here("data", "dataPrep", "removed_periods.feather"))
```

```{r remove-capture-cage-getroosts}
#| eval: false
# Identify and remove capture dates using Marta's code --------------------
roosts0 <- get_roosts_df(df = removed_periods, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
save(roosts0, file = here::here("data", "dataPrep", "roosts0.Rda"))

@@

```{r}
# Load information about the capture sites
capture_sites <- read.csv(here::here("data", "raw", "capture_sites.csv"))
load(here::here("data", "dataPrep", "roosts0.Rda"))
AllCarmelDates <- read.csv(here::here("data", "raw", "all_captures_carmel_2010-2021.csv"))
```

@@

write_feather(removed_captures, here::here("data", "dataPrep", "removed_captures.feather"))

@@

```{r attach-age-sex}
#| message: false
removed_captures <- read_feather(here::here("data", "dataPrep", "removed_captures.feather"))
age_sex <- read_excel(here::here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "all gps tags")[,1:35] %>%
  dplyr::select(Nili_id, birth_year, sex) %>%
  distinct()
```

@@

write_feather(with_age_sex, here::here("data", "dataPrep", "with_age_sex.feather"))

@@

```{r}
mask <- sf::st_read(here::here("data", "raw", "CutOffRegion.kml"))
with_age_sex <- read_feather(here::here("data", "dataPrep", "with_age_sex.feather"))
```

@@

save(cleaned, file = here::here("data", "dataPrep", "cleaned.Rda"))

@@

# Separate the seasons -----------------------------
seasons_list <- with_seasons %>%
  group_by(seasonUnique) %>%
  group_split(.keep = T)
length(seasons_list) # should be 9
# extract the season names in case we need a separate vector at some point.
season_names <- map_chr(seasons_list, ~as.character(.x$seasonUnique[1])) # ok good, these are in the right order
length(season_names) # should also be 9
save(seasons_list, file = here::here("data", "dataPrep", "seasons_list.Rda"))
save(season_names, file = here::here("data", "dataPrep", "season_names.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "seasons_list.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

@@

save(low_fix_rate_indivs, 
     file = here::here("data", "dataPrep", "low_fix_rate_indivs.Rda"))
# Don't need to save with_modes because we've already used this step to identify low fix rate individuals.

@@

```{r}
load(here::here("data", "dataPrep", "seasons_list.Rda")) 
load(here::here("data", "dataPrep", "low_fix_rate_indivs.Rda"))
```

```{r}
removed_lfr <- map2(seasons_list, low_fix_rate_indivs,                     ~.x %>% filter(!(Nili_id %in% .y))) 
save(removed_lfr, file = here::here("data", "dataPrep", "removed_lfr.Rda"))  
# tidy up 
rm(seasons_list) 
rm(low_fix_rate_indivs) 
rm(removed_lfr) 
gc()
```

@@

```{r}
#| eval: false
# Get roosts for each season ----------------------------------------------
roosts <- purrr::map(removed_lfr, ~vultureUtils::get_roosts_df(df = .x, id = "Nili_id"))
roosts <- roosts %>%
  map(., ~st_as_sf(.x, crs = "WGS84", 
                   coords = c("location_long", "location_lat"), 
                   remove = F))
save(roosts, file = here::here("data", "dataPrep", "roosts.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "removed_lfr.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

@@

save(removed_nighttime, file = here::here("data", "dataPrep", "removed_nighttime.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "removed_nighttime.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

@@

save(removed_northern, file = here::here("data", "dataPrep", "removed_northern.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "removed_northern.Rda"))
```

@@

save(removed_lowppd, file = here::here("data", "dataPrep", "removed_lowppd.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "removed_lowppd.Rda"))
load(here::here("data", "dataPrep", "season_names.Rda"))
```

@@

save(removed_too_few_days, file = here::here("data", "dataPrep", "removed_too_few_days.Rda"))
save(season_durations, file = here::here("data", "dataPrep", "season_durations.Rda"))

@@

```{r}
load(here::here("data", "dataPrep", "removed_too_few_days.Rda"))
```

```{r}
#| eval: false
# Get elevation rasters ---------------------------------------------------
sf <- map(removed_too_few_days, ~st_as_sf(.x, coords = c("location_lat", "location_long"), remove = F, crs = "WGS84"))
elevs_z10 <- map(sf, ~elevatr::get_elev_raster(.x , z = 10))
save(elevs_z10, file = here::here("data", "dataPrep", "elevs_z10.Rda"))
save(sf, file = here::here("data", "dataPrep", "sf.Rda"))
```

```{r}
load(here::here("data", "dataPrep", "elevs_z10.Rda"))
load(here::here("data", "dataPrep", "sf.Rda"))
```

@@

save(with_altitudes, file = here::here("data", "dataPrep", "with_altitudes.Rda"))

### 2024-02-14 scripts/00.0_dataPrep_copy.qmd
fixed file names to use here::here()

```{r}
# joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
# year22 <- joined %>%
#   filter(lubridate::year(dateOnly) == "2022")
write_csv(year22, file = here::here("data", "dataPrep", "year22.csv"))
```

@@

```{r load-hospital}
year22 <- read_csv("data/dataPrep/year22.csv")
periods_to_remove <- read_excel("data/raw/whoswho_vultures_20230920_new.xlsx", sheet = "periods_to_remove")
```

@@

# Actually REMOVE the periods to remove...
removed_periods <- removal_annotated %>%
  filter(status == "valid")
nrow(removed_periods)
length(unique(removal_annotated$Nili_id))
nrow(year22) == nrow(removal_annotated) # true
nrow(removed_periods) != nrow(removal_annotated) # true

>>>>

```{r}
# joined <- read_feather(here::here("data", "dataPrep", "joined.feather"))
# year22 <- joined %>%
#   filter(lubridate::year(dateOnly) == "2022")
# write_csv(year22, file = here::here("data", "dataPrep", "year22.csv"))
```

@@

```{r load-hospital}
year22 <- read_csv(here("data", "dataPrep", "year22.csv"))
periods_to_remove <- read_excel(here("data", "raw", "whoswho_vultures_20230920_new.xlsx"), sheet = "periods_to_remove")
```

@@

# Actually REMOVE the periods to remove...
removed_periods <- removal_annotated %>%
  filter(status == "valid")
nrow(removed_periods)
length(unique(removal_annotated$Nili_id))
